#Load libraries
```{r}
#libraries
library(tidyverse)
library(here)
library(ggmap)
library(ggpmisc)
library(stringr)
library(sf)
library(sp)
library(spatialEco)
library(rnaturalearth)
library(taxize)
#check here
here::here()

my_theme <- function() {theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(size=12, face= "bold", colour= "black"), axis.title.x = element_text(size=12, face="bold", colour = "black"), axis.title.y = element_text(size=12, face="bold", colour = "black"))}
```

#Get clean data
```{r}
fishdf <- read.csv(here("data", "processeddata", "RFS_AbundanceBiomassDepth.csv"))
fishdf$date <- as.Date(fishdf$date)
```

#Post cleaning
We are interested in how the reef fish community has changed over time.

Variables we could include: 
 - Location (Lat/Long)
 - Date
 - Family
 - Length
 - Weight
 - Sex
 - Maturity
 - Depth
 - Temperature (surface/bottom)
 - Salinity (surface/bottom)
 - Dissolved Oxygen (surface/bottom)

We could do a multidimensional analysis of community composition, a tree based model of what environmental factors drive distributions of fishes, model of commerically important species. Or all of these :)

#Map
```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
world <- st_make_valid(world)
sab <- st_crop(world, xmin = -82, xmax = -76, ymin = 26, ymax = 36)

ggplot(data = sab) + geom_sf() + geom_point(data = fishdf, mapping = aes(x = longitudestart, y = latitudestart)) + labs(title = "All Abundance/Biomass Locations", x = "Longitude", y = "Latitude") + my_theme() + facet_wrap(~year)
```
Definitely some changes in sampling effort in 2010 (from website). Let's only use data after 2010. 

```{r}
fishdfpost2010 <- fishdf %>% dplyr::filter(year > 2009)

ggplot(data = sab) + geom_sf() + geom_point(data = fishdfpost2010, mapping = aes(x = longitudestart, y = latitudestart)) + labs(title = "All Abundance/Biomass Locations", x = "Longitude", y = "Latitude") + my_theme() + facet_wrap(~year)
```

#Sampling effort
```{r}
#sampling effort EDA
ggplot(data = fishdfpost2010, mapping = aes(x = year, y = month)) + geom_point()

#seems like early and late months maybe shouldn't be sampled
fishdfpost2010 %>% group_by(month) %>% ggplot(mapping = aes(x = year, y = speciestotalweight)) + geom_point() + geom_line() + facet_wrap(~month)

fishdfpost2010 <- filter(fishdfpost2010, month %in% 5:9)
#temperature
ggplot(data = fishdfpost2010, mapping = aes(x = date, y = tempsurface)) + geom_point() + geom_smooth(method = "lm")

#remove temps at 0
fishdfpost2010 <- filter(fishdfpost2010, tempsurface != 0)

#temperature again
ggplot(data = fishdfpost2010, mapping = aes(x = date, y = tempsurface)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4)


floor_decade = function(value){ return(value - value %% 5) }
decade <- mutate(fishdfpost2010, decade = floor_decade(year))

ggplot(data = sab) + geom_sf() + geom_point(data = decade , aes(y = latitudestart, x = longitudestart), size = 0.5) + xlab("Longitude") + ylab("Latitude") +  theme(legend.position = "none") + facet_wrap(~decade) + ggtitle("Sampling Effort over Time") + my_theme()

# ggsave(here("figures", "SamplingEffortby5yrs.png"), width = 6, height = 5, unit = "in")
```

Sampling effort seems fine, but I would prefer to have station IDs. 

#Abundance and Biomass
```{r}
#Let's start with abundances and biomass trends
annualmetrics <- fishdfpost2010 %>% group_by(speciescommonname, year) %>% summarize(abundance = sum(numbertotal, na.rm = TRUE), biomass = sum(speciestotalweight, na.rm = TRUE), meansurftemp = mean(tempsurface, na.rm = TRUE), meansurfsal = mean(salinitysurface, na.rm = TRUE), meansurfdo = mean(sdo, na.rm = TRUE), meanlatitude = mean(latitudestart, na.rm = TRUE))

metrics <- fishdfpost2010 %>% group_by(speciescommonname) %>% summarize(abundance = sum(numbertotal, na.rm = TRUE), biomass = sum(speciestotalweight, na.rm = TRUE), meansurftemp = mean(tempsurface, na.rm = TRUE), meansurfsal = mean(salinitysurface, na.rm = TRUE), meansurfdo = mean(sdo, na.rm = TRUE))

arrange(metrics, desc(abundance))

#top 10
top10 <- head(arrange(metrics, desc(abundance)), n = 10)
topnames <- top10$speciescommonname
topdata <- filter(annualmetrics, speciescommonname %in% topnames)

#abundance top 
ggplot(data = topdata, mapping = aes(x = year, y = abundance, color = speciescommonname)) + geom_point() + geom_line() + ggtitle("Top 10 Abundance Trends Reef Survey") + my_theme() + labs(x = "Year", y = "Abundance", color = "Species Common Name")
ggsave(here("figures", "AbundanceTrendsTop10.png"), width = 6, height = 5, unit = "in")

#biomass top 10
ggplot(data = topdata, mapping = aes(x = year, y = biomass, color = speciescommonname)) + geom_point() + geom_line() + ggtitle("Top 10 Biomass Trends Reef Survey") + my_theme() + labs(x = "Year", y = "Biomass", color = "Species Common Name") 
ggsave(here("figures", "BiomassTrendsTop10.png"), width = 6, height = 5, unit = "in")

#top 10 map
topdataall <- filter(fishdfpost2010, speciescommonname %in% topnames)
ggplot(data = sab) + geom_sf() + geom_point(data = topdataall, aes(y = latitudestart, x = longitudestart, color = speciescommonname)) + xlab("Longitude") + ylab("Latitude") 

#families
families <- fishdfpost2010 %>% group_by(family) %>% summarize(abundance = sum(numbertotal, na.rm = TRUE), biomass = sum(speciestotalweight, na.rm = TRUE))
arrange(families, desc(abundance))

#one species
triggerfish <- filter(fishdfpost2010, speciescommonname == "GRAYTRIGGERFISH")
ggplot(data = sab) + geom_sf()  + geom_point(data = triggerfish, aes(y = latitudestart, x = longitudestart, color = year), size = 0.5) + xlab("Longitude") + ylab("Latitude") 

```

# Environmental Variables
```{r}
#Temperature
fishdfpost2010 %>% ggplot(aes(x = date, y = tempsurface)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = " Changes in Surface Temperature over Time") + xlab("Date") + ylab("Temperature (\u00B0C)") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

fishdfpost2010 %>% ggplot(aes(x = date, y = tempbottom)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = " Changes in Bottom Temperature over Time") + xlab("Date") + ylab("Temperature (\u00B0C)") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#by top 10 species
topdataall %>% ggplot(aes(x = date, y = tempsurface)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = " Changes in Temperature over Time") + facet_wrap(~speciescommonname)  + xlab("Date") + ylab("Temperature (\u00B0C)") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

topdata %>% ggplot(aes(x = meansurftemp, y = abundance)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = " Changes in Temperature with Abundance") + facet_wrap(~speciescommonname)  + xlab("Temperature (\u00B0C)") + ylab("Abundance") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

topdata %>% ggplot(aes(x = meansurftemp, y = biomass)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = " Changes in Temperature with Biomass") + facet_wrap(~speciescommonname)  + xlab("Temperature (\u00B0C)") + ylab("Biomass ") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#Salinity
#these look weird
fishdfpost2010 %>% filter(salinitysurface > 25) %>% ggplot(aes(x = date, y = salinitysurface)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = "Changes in Surface Salinity over Time") + xlab("Date") + ylab("ppt") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

fishdfpost2010 %>% filter(salinitysurface > 25) %>% ggplot(aes(x = date, y = salinitybottom)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = "Changes in Bottom Salinity over Time") + xlab("Date") + ylab("ppt") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#Take out 0 salinities
fishdfpost2010 <- fishdfpost2010 %>% filter(salinitysurface != 0)
fishdfpost2010 <- fishdfpost2010 %>% filter(salinitybottom != 0)

#Dissolved Oxygen
#these look weird
fishdfpost2010  %>% ggplot(aes(x = date, y = sdo)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = "Changes in Surface Dissolved Oxygen over Time") + xlab("Date") + ylab("mg/dL") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
#Take out 0 sdo
fishdfpost2010 <- fishdfpost2010 %>% filter(sdo > 0)

fishdfpost2010  %>% ggplot(aes(x = date, y = bdo)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = "Changes in Bottom Dissolved Oxygen over Time") + xlab("Date") + ylab("mg/dL") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
#Take out 0 bdo
fishdfpost2010 <- fishdfpost2010 %>% filter(bdo > 0)
```

#Write final cleaned
```{r}
write.csv(fishdfpost2010, here("data", "processeddata", "RFS_AbundanceBiomassDepth.csv"))
```

#Merge with specimen and length data
We just want total length, so we'll select len1
```{r}
specimen <- read.csv(here("data", "processeddata", "RFS_Specimen.csv"))
specimen <- unique(specimen)
specimen <- specimen %>% dplyr::select(eventname, date, year, month, speciescommonname, speciesscientificname, duration, len1, weight, sexdescription, maturitydescription, diseasedesc)
fishdfpost2010 <- fishdfpost2010 %>% dplyr::select(eventname, date, year, month, eventname, family, speciesscientificname, speciescommonname, numbertotal, speciestotalweight, tempsurface, tempbottom, salinitysurface, salinitybottom, sdo, bdo, latitudestart, latitudeend, depthstart, duration)

fishdfpost2010merged <- merge(fishdfpost2010, specimen, by = c("eventname", "date", "year", "month", "speciesscientificname"))

# write.csv(fishdfpost2010merged, here("data", "processeddata", "RFS_AbundanceBiomassDepthLength.csv"))

#This is the same length information I don't want it
# length <- read.csv(here("data", "processeddata", "RFS_Length.csv"))
# length <- length %>% dplyr::select(eventname, date, year, month, speciescommonname, speciesscientificname, length)
# fishdfpost2010merged <- merge(fishdfpost2010merged, length, by = c("eventname", "date", "year", "month", "speciesscientificname"))


```