#Load libraries
```{r}
#libraries
library(tidyverse)
library(here)
library(ggmap)
library(ggpmisc)
library(stringr)
library(sf)
library(sp)
library(spatialEco)
# devtools::install_github("jeffreyevans/spatialEco")
library(rnaturalearth)
library(taxize)
#check here
here::here()

my_theme <- function() {theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(size=12, face= "bold", colour= "black"), axis.title.x = element_text(size=12, face="bold", colour = "black"), axis.title.y = element_text(size=12, face="bold", colour = "black"))}
```

#Get clean data
```{r}
fishdf <- read.csv(here("data", "processeddata", "Clean_SEUS.csv"))
fishdf$date <- as.Date(fishdf$date)
```

#Post cleaning
We are interested in how the reef fish community has changed over time.

Variables we could include: 
 - Location (Lat/Long)
 - Date
 - Family
 - Weight
 - Depth
 - Temperature (surface/bottom)

We could do a multidimensional analysis of community composition, a tree based model of what environmental factors drive distributions of fishes, model of commercially important species. Or all of these :)

#Map
```{r}
library(sf)
library(sp)
library(spatialEco)
# devtools::install_github("jeffreyevans/spatialEco")
library(rnaturalearth)

world <- ne_countries(scale = "large", returnclass = "sf")
world <- st_make_valid(world)
sab <- st_crop(world, xmin = -82, xmax = -76, ymin = 26, ymax = 36)

ggplot(data = sab) + geom_sf() + geom_point(data = fishdf, mapping = aes(x = longitude, y = latitude)) + labs(title = "All Abundance/Biomass Locations", x = "Longitude", y = "Latitude") + my_theme() 

floor_decade = function(value){ return(value - value %% 5) }
decade <- mutate(fishdf, decade = floor_decade(year))

ggplot(data = sab) + geom_sf() + geom_point(data = decade , aes(y = latitude, x = longitude), size = 0.5) + xlab("Longitude") + ylab("Latitude") +  theme(legend.position = "none") + facet_grid(~decade) + ggtitle("Pre-Data Cleaning Sampling Effort over Time") + my_theme() + scale_x_continuous(breaks=seq(-80, -77, 3))

ggsave(here("figures", "PreCleanedSamplingEffortby5yrs.png"), width = 6, height = 5, unit = "in")

ggplot(fishdf, mapping = aes(x = year, y = station)) + geom_point(size = 0.5, alpha = 0.5)


#Definitely some changes in sampling effort over time. Let's only use resampled stations and do just after 2000. 

resampled <- fishdf  %>%
  group_by(station) %>%
  filter(all(2000:2019 %in% year)) %>% filter(year > 1999)

ggplot(resampled, mapping = aes(x = year, y = station)) + geom_point(size = 0.5, alpha = 0.5)

ggplot(data = sab) + geom_sf() + geom_point(data = resampled, mapping = aes(x = longitude, y = latitude)) + labs(title = "All Resampled Abundance/Biomass Locations", x = "Longitude", y = "Latitude") + my_theme() 

decade <- mutate(resampled, decade = floor_decade(year))

ggplot(data = sab) + geom_sf() + geom_point(data = decade , aes(y = latitude, x = longitude), size = 0.5) + xlab("Longitude") + ylab("Latitude") +  theme(legend.position = "none") + facet_grid(~decade) + ggtitle("Post-Cleaning Sampling Effort over Time") + my_theme() + scale_x_continuous(breaks=seq(-80, -77, 3)) + scale_x_continuous(breaks=seq(-80, -77, 3))

ggsave(here("figures", "PostCleanedSamplingEffortby5yrs.png"), width = 6, height = 5, unit = "in")
```

#Sampling effort
```{r}

#seems like months aren't surveyed equally. Should be fine
ggplot(resampled, mapping = aes(x = year, y = station)) + geom_point(size = 0.5, alpha = 0.5) + facet_wrap(~month)

#temperature
ggplot(data = resampled, mapping = aes(x = date, y = sst)) + geom_point() + geom_smooth(method = "lm")

#temperature again
ggplot(data = resampled, mapping = aes(x = date, y = sst)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4)


floor_decade = function(value){ return(value - value %% 5) }
decade <- mutate(resampled, decade = floor_decade(year))

ggplot(data = sab) + geom_sf() + geom_point(data = decade , aes(y = latitude, x = longitude), size = 0.5) + xlab("Longitude") + ylab("Latitude") +  theme(legend.position = "none") + facet_wrap(~decade) + ggtitle("Sampling Effort over Time") + my_theme()

ggsave(here("figures", "SamplingEffortby5yrs.png"), width = 6, height = 5, unit = "in")
```

Sampling effort seems good!

#Bathymetry of sites
```{r}
library(ncdf4)
library(lubridate)
library(rerddap)
library(doParallel)
library(marmap)
library(geotools)
library(geosphere)

xlims <- c(-82,-76)
ylims <- c(26,36)
bathydata<- getNOAA.bathy(xlims[1], xlims[2], ylims[1], ylims[2])

autoplot.bathy(bathydata, geom=c("r", "c"),coast="TRUE") + 
  scale_fill_etopo() +
  ylim(26,36) +
  xlim(-82,-76)+
  geom_point(data=resampled, aes(x=longitude, y=latitude),  alpha=1)  + my_theme() + labs(title = "SEUS Bathymetry Data", x = "Longitude", y = "Latitude", fill = "Depth") + guides(color="none")+ scale_fill_gradient2(low="darkblue", mid="lightblue", high="white", na.value = "black") 

ggsave(here("figures", "BathymetryMap.png"), width = 6, height = 5, unit = "in")
```

#Abundance and Biomass
```{r}
#Let's start with abundances and biomass trends
annualmetrics <- resampled %>% group_by(accepted_name, year) %>% summarize(abundance = sum(num_cpue, na.rm = TRUE), biomass = sum(wgt_cpue, na.rm = TRUE), meansurftemp = mean(sst, na.rm = TRUE), meanlatitude = mean(latitude, na.rm = TRUE))

metrics <- resampled %>% group_by(accepted_name) %>% summarize(abundance = sum(num_cpue, na.rm = TRUE), biomass = sum(wgt_cpue, na.rm = TRUE), meansurftemp = mean(sst, na.rm = TRUE), meanlatitude = mean(latitude, na.rm = TRUE))

arrange(metrics, desc(abundance))

#top 10
top10 <- head(arrange(metrics, desc(abundance)), n = 10)
topnames <- top10$accepted_name
topdata <- filter(annualmetrics, accepted_name %in% topnames)

#colors :)
palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733",
                               "#44AA99", "#999933", "#332288", "#AA4499", "#882255", "#E66101","#6699CC","#E7298A", "#CA0020",  "#525252", "#661100",  "#41B6C4", "#999999")

#abundance top 
ggplot(data = topdata, mapping = aes(x = year, y = abundance, color = accepted_name)) + geom_point() + geom_line() + ggtitle("Top 10 Species Annual Abundance Trends SEUS") + my_theme() + labs(x = "Year", y = "Abundance", color = "Species Scientific Name") + scale_color_manual(values = palette)
ggsave(here("figures", "AbundanceTrendsTop10.png"), width = 6, height = 5, unit = "in")

#biomass top 10
ggplot(data = topdata, mapping = aes(x = year, y = biomass, color = accepted_name)) + geom_point() + geom_line() + ggtitle("Top 10 Species Annual Biomass Trends SEUS") + my_theme() + labs(x = "Year", y = "Biomass", color = "Species Scientific Name") + scale_color_manual(values = palette)
ggsave(here("figures", "BiomassTrendsTop10.png"), width = 6, height = 5, unit = "in")

#top 10 map
topdataall <- filter(resampled, accepted_name %in% topnames)
ggplot(data = sab) + geom_sf() + geom_point(data = topdataall, aes(y = latitude, x = longitude, color = accepted_name)) + xlab("Longitude") + ylab("Latitude") 

#families
families <- resampled %>% group_by(family) %>% summarize(abundance = sum(num_cpue, na.rm = TRUE), biomass = sum(wgt_cpue, na.rm = TRUE))
arrange(families, desc(abundance))

#one species
croaker <- filter(resampled, accepted_name == "Micropogonias undulatus")
ggplot(data = sab) + geom_sf()  + geom_point(data = croaker, aes(y = latitude, x = longitude, color = year)) + xlab("Longitude") + ylab("Latitude") 

```

# Environmental Variables
```{r}
#Temperature
resampled %>% ggplot(aes(x = date, y = sst)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = " Changes in Surface Temperature over Time") + xlab("Date") + ylab("Temperature (\u00B0C)") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

resampled %>% ggplot(aes(x = date, y = sbt)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = " Changes in Bottom Temperature over Time") + xlab("Date") + ylab("Temperature (\u00B0C)") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#by top 10 species
topdataall %>% ggplot(aes(x = date, y = sst)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = " Changes in Temperature over Time") + facet_wrap(~accepted_name)  + xlab("Date") + ylab("Temperature (\u00B0C)") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

topdata %>% ggplot(aes(x = meansurftemp, y = abundance)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = " Changes in Temperature with Abundance") + facet_wrap(~accepted_name)  + xlab("Temperature (\u00B0C)") + ylab("Abundance") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

topdata %>% ggplot(aes(x = meansurftemp, y = biomass)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = " Changes in Temperature with Biomass") + facet_wrap(~accepted_name)  + xlab("Temperature (\u00B0C)") + ylab("Biomass ") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

#Write final cleaned
```{r}
write.csv(resampled, here("data", "processeddata", "Final_SEUS.csv"))
```
