#Load libraries
```{r}
library(tidyverse)
library(MASS)
library(here)
#check here
here::here()
library(gjam)
my_theme <- function() {theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(size=12, face= "bold", colour= "black"), axis.title.x = element_text(size=12, face="bold", colour = "black"), axis.title.y = element_text(size=12, face="bold", colour = "black"))}
```

#Get clean data
```{r}
fishdf <- read.csv(here("data", "processeddata", "Final_SEUS.csv"))
fishdf$date <- as.Date(fishdf$date)
```

We are interested in how the reef fish community has changed over time and want to build a tree based model of what environmental factors drive distributions of fishes

Variables we could include: 
 - Location (Lat/Long)
 - Date
 - Family
 - Weight
 - Depth
 - Temperature (surface/bottom)
 
#Select variables of interest
```{r}
names(fishdf)
#there are significantly fewer rows with maturity descriptions 
fishdf <- fishdf %>% dplyr::filter(station == "29M4") %>% dplyr::select(date, year, month, day, accepted_name, family, station, latitude, longitude, depth, sbt, sst, num_cpue, wgt_cpue)
```

#Xdata (explanatory) and ydata (response)
```{r}
xdata <- fishdf %>% dplyr::select(year, month, day, depth, sbt, sst)
ydata_long <- fishdf %>% dplyr::select(accepted_name, num_cpue, station, date)
ydata <- ydata_long %>% group_by(accepted_name) %>% mutate(row = row_number()) %>% pivot_wider(names_from = accepted_name, values_from = num_cpue) %>% dplyr::select(-c(station, date, row))
ydata[is.na(ydata)] <- 0
```

#Start model
```{r}
ml <- list(ng = 1000, burnin = 500, typeNames = 'CA')
out <- gjam(formula = ~year + month + day + depth + sbt + sst, xdata = xdata, ydata = ydata, modelList = ml)

save(out1, file='out_SEUS.Rdata')
plotPars <- list(PLOTALLY=T, SAVEPLOTS = T, SMALLPLOTS = T, GRIDPLOTS = T, outFolder = 'out1')
gjamPlot(out1, plotPars=plotPars)
```