#Load libraries
```{r}
library(tidyverse)
library(here)
library(ggmap)
library(ggpmisc)
library(stringr)
library(sf)
library(sp)
library(spatialEco)
library(rnaturalearth)
library(taxize)
#check here
here::here()

library(tidymodels)
library(tidyflow)
library(rpart.plot)
library(vip)
library(baguette)
library(ranger)

my_theme <- function() {theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(size=12, face= "bold", colour= "black"), axis.title.x = element_text(size=12, face="bold", colour = "black"), axis.title.y = element_text(size=12, face="bold", colour = "black"))}
```

#Get clean data
```{r}
fishdf <- read.csv(here("data", "processeddata", "RFS_AbundanceBiomassDepth.csv"))
fishdf$date <- as.Date(fishdf$date)
```

We are interested in how the reef fish community has changed over time and want to build a tree based model of what environmental factors drive distributions of fishes

Variables we could include: 
 - Location (Lat/Long)
 - Date
 - Family
 - Length
 - Weight
 - Sex
 - Maturity
 - Depth
 - Temperature (surface/bottom)
 - Salinity (surface/bottom)
 - Dissolved Oxygen (surface/bottom)
 
#Select variables of interest
```{r}
fishdf <- fishdf %>% dplyr::select(date, year, month, speciesscientificname, speciestotalweight, tempsurface, tempbottom, salinitysurface, salinitybottom, sdo, bdo, latitudestart, longitudestart, depthstart, family)
```
 
 