#Load libraries
```{r}
library(tidyverse)
library(here)
library(ggmap)
library(ggpmisc)
library(stringr)
library(sf)
library(sp)
library(spatialEco)
library(rnaturalearth)
library(taxize)
#check here
here::here()

library(tidymodels)
library(tidyflow)
library(rpart.plot)
library(vip)
library(baguette)
library(ranger)

my_theme <- function() {theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(size=12, face= "bold", colour= "black"), axis.title.x = element_text(size=12, face="bold", colour = "black"), axis.title.y = element_text(size=12, face="bold", colour = "black"))}
```

#Get clean data
```{r}
fishdf <- read.csv(here("data", "processeddata", "Final_SEUS.csv"))
fishdf$date <- as.Date(fishdf$date)
```

We are interested in how the fish community has changed over time and want to build a tree based model of what environmental factors drive distributions of fishes

Variables we could include: 
 - Location (Lat/Long)
 - Date
 - Family
 - Weight
 - Depth
 - Temperature (surface/bottom)
 
#Select variables of interest
```{r}
names(fishdf)
#there are significantly fewer rows with maturity descriptions 
fishdf <- fishdf %>% dplyr::select(date, year, month, accepted_name, family, station, latitude, longitude, depth, sbt, sst, num_cpue, wgt_cpue)
```

#Pick out species of interest
```{r}
annualmetrics <- fishdf %>% group_by(accepted_name, year) %>% summarize(abundance = sum(num_cpue, na.rm = TRUE), biomass = sum(wgt_cpue, na.rm = TRUE), meansurftemp = mean(sst, na.rm = TRUE), meanlatitude = mean(latitude, na.rm = TRUE))

metrics <- fishdf %>% group_by(accepted_name) %>% summarize(abundance = sum(num_cpue, na.rm = TRUE), biomass = sum(wgt_cpue, na.rm = TRUE), meansurftemp = mean(sst, na.rm = TRUE), meanlatitude = mean(latitude, na.rm = TRUE))

arrange(metrics, desc(abundance))


top10 <- metrics %>%                                      
  arrange(desc(abundance)) %>% 
  group_by(accepted_name) %>% ungroup() %>%
  slice(1:10)


top10names <- top10$accepted_name
```

#Model building
Lets say we want to know what factors determine croaker presence
```{r}
croaker <- fishdf %>% dplyr::filter(accepted_name == "Micropogonias undulatus") %>% dplyr::select(year, month, latitude, longitude, depth, sbt, sst, wgt_cpue) %>% na.omit()
```
#Basic models
```{r}
# Define the decision tree and tell it the the dependent
# variable is continuous ('mode' = 'regression')
mod1 <- set_engine(decision_tree(mode = "regression"), "rpart")

tflow <-
  # Plug the data
  croaker %>%
  # Begin the tidyflow
  tidyflow(seed = 23151) %>%
  # Separate the data into training/testing (we are keeping 3/4 of the data for training)
  plug_split(initial_split, prop = 3/4) %>%
  # Plug the formula
  plug_formula(wgt_cpue ~ .) %>%
  # Plug the model
  plug_model(mod1)

vanilla_fit <- fit(tflow)
tree <- pull_tflow_fit(vanilla_fit)$fit
rpart.plot(tree)

vanilla_fit %>%
  pull_tflow_fit() %>%
  .[['fit']] %>%
  vip() +
  theme_minimal() + labs(title = "Atlantic Croaker Decision Tree Variable Importances", x = "Variable", y = "Importance")
```
Fish are determined by time and space. 

#Random forests
```{r}
# Define the random forest
rf_mod <-
  rand_forest(mode = "regression") %>%
  set_engine("ranger", importance = "impurity")

# Define the `tidyflow` with the random forest model
# and include all variables (including scie_score and read_score)
tflow <-
  croaker %>%
  tidyflow(seed = 23151) %>%
  plug_formula(wgt_cpue ~ .) %>%
  plug_split(initial_split) %>%
  plug_model(rf_mod)

res_rf <- tflow %>% fit()

res_rf %>%
  predict_training() %>%
  rmse(wgt_cpue, .pred)

res_rf %>%
  pull_tflow_fit() %>%
  .[['fit']] %>%
  vip() +
  theme_minimal() + labs(title = "Atlantic Croaker Random Forest Variable Importances", x = "Variable", y = "Importance")
```
Biggest drivers are temperature and latitude (basically temperature)

#Tune random forests
I am going to mess with the "mtry" argument to 1/3 of the amount of variables (3), i.e. the number of columns to be used in each split
```{r}
rf_mod <-
  rand_forest(mode = "regression", mtry = 3) %>%
  set_engine("ranger")

res_rf <- tflow %>% replace_model(rf_mod) %>% fit()

res_rf %>%
  predict_training() %>%
  rmse(wgt_cpue, .pred)

#Predict on testing
rf_rmse <-
  res_rf %>%
  predict_testing() %>%
  rmse(wgt_cpue, .pred) %>%
  pull(.estimate)
```
RMSE is slightly smaller, with error of 97 kg/min

Let's use tidyflow's grid_regular to obtain an evenly spaced set of values for trees -- we'll divide the data into training/testing, specify a cross-validation and plug in the random forest and a grid_regular
```{r}
rf_mod <-
  rand_forest(mode = "regression",
              trees = tune()) %>%
  set_engine("ranger")

tflow <-
  croaker %>% 
  tidyflow(seed = 2151) %>%
  plug_split(initial_split) %>%
  plug_resample(vfold_cv) %>%
  plug_grid(grid_regular, levels = 10) %>%
  plug_formula(wgt_cpue ~ .) %>%
  plug_model(rf_mod)

tflow

res <- tflow %>% fit()
res %>% pull_tflow_fit_tuning() %>% autoplot()

final_mod <- res %>% complete_tflow(metric = "rmse")

final_mod %>%
  predict_training() %>%
  rmse(wgt_cpue, .pred)
```
RMSE is slightly better.

#xgboost
```{r}
boost_mod <-
  boost_tree(mode = "regression", trees = 500) %>%
  set_engine("xgboost")

tflow <-
  croaker %>%
  tidyflow(seed = 51231) %>%
  plug_formula(wgt_cpue ~ .) %>%
  plug_split(initial_split) %>%
  plug_model(boost_mod)

res_boost <- fit(tflow)

rmse_gb_train <-
  res_boost %>%
  predict_training() %>%
  rmse(wgt_cpue, .pred)

rmse_gb_train

```
Pretty good improvement in RMSE!

#Compare RF and xgboost on testing
```{r}
gb_rmse <-
  res_boost %>%
  predict_testing() %>%
  rmse(wgt_cpue, .pred) %>%
  pull(.estimate)

c(
  "Random Forest" = rf_rmse,
  "Extreme Gradient Boosting" = gb_rmse)
```
RF performs a little better. Interesting! Let's pick out a the top 10 species and run xgboost anyway to figure out variable importance 

#Multiple species
```{r}
rmse <- list()
vips <- list()
allnames <- unique(fishdf$accepted_name)
namestokeep <- list()
fishdf %>% group_by(accepted_name) %>% summarize(count = n())
for (i in 1:length(allnames)) {
  label <- str_to_title(allnames[i])
  data <- filter(fishdf, accepted_name == allnames[i])
  data <- data %>% dplyr::select(year, month, latitude, longitude, depth, sbt, sst, wgt_cpue) %>% na.omit()
    if(nrow(data) > 10) {
      namestokeep <-  c(namestokeep, allnames[i])
        
      boost_mod <-
      boost_tree(mode = "regression", trees = 500) %>%
      set_engine("xgboost")
    
      tflow <-
        data %>%
        tidyflow(seed = 51231) %>%
        plug_formula(wgt_cpue ~ .) %>%
        plug_split(initial_split) %>%
        plug_model(boost_mod)
    
      res_boost <- fit(tflow)
    
      rmse_gb_test <-
        res_boost %>%
        predict_testing() %>%
        rmse(wgt_cpue, .pred) %>%
      pull(.estimate)
    
      rmse[i] <- rmse_gb_test
      
     vip <-  res_boost %>%
        pull_tflow_fit() %>%
        vip()
     vip <- vip[["data"]]
     vip$Species <- label
     vips[[i]] <- vip
    
      p <- res_boost %>%
        pull_tflow_fit() %>%
        .[['fit']] %>%
        vip() +
        theme_minimal() + labs(title = paste0(label, " xgboost Variable Importances"), x = "Variable", y = "Importance") + my_theme()
      print(p)
      # ggsave(here("figures", paste0("VariableImportance", label, ".png")), width = 6, height = 4, unit = "in")
    }else{
      print(paste("Not enough data for", allnames[i]))
    }
  
}

finalrmses <- data.frame(unlist(namestokeep), unlist(rmse))
finalvips <- do.call(rbind, vips)

finalvips2 <- finalvips %>% group_by(Species) %>% dplyr::filter(Importance == max(Importance))
ggplot(finalvips2, mapping = aes(x = Variable)) + geom_bar() + my_theme() + labs(title = "xgboost Highest Variable Importance within each Species Model", x = "Variable", y = "Number of Species with Variable as Important")
ggsave(here("figures", "SpeciesCountofVariableImportance.png"), width = 6, height = 4, unit = "in")
```

#By family
```{r}
families <- unique(fishdf$family)
rmse <- list()
vips <- list()
namestosave <- list()
    
for (i in 1:length(families)) {
  label <- str_to_title(families[i])
  data <- filter(fishdf, family == families[i])
  data <- data %>% dplyr::select(year, month, latitude, longitude, depth, sbt, sst, wgt_cpue) %>% na.omit()
  if(nrow(data) > 10) {
  namestosave <- c(namestosave, families[i])
  boost_mod <-
  boost_tree(mode = "regression", trees = 500) %>%
  set_engine("xgboost")

  tflow <-
    data %>%
    tidyflow(seed = 51231) %>%
    plug_formula(wgt_cpue ~ .) %>%
    plug_split(initial_split) %>%
    plug_model(boost_mod)

  res_boost <- fit(tflow)

  rmse_gb_test <-
    res_boost %>%
    predict_testing() %>%
    rmse(wgt_cpue, .pred) %>%
  pull(.estimate)

  rmse[i] <- rmse_gb_test
  
 vip <-  res_boost %>%
    pull_tflow_fit() %>%
    vip()
 vip <- vip[["data"]]
 vip$Family <- label
 vips[[i]] <- vip

  p <- res_boost %>%
    pull_tflow_fit() %>%
    .[['fit']] %>%
    vip() +
    theme_minimal() + labs(title = paste0(label, " xgboost Variable Importances"), x = "Variable", y = "Importance") + my_theme()
  print(p)
  ggsave(here("figures", paste0("VariableImportance", label, ".png")), width = 6, height = 4, unit = "in")
  }
  
}

finalrmses <- data.frame(unlist(namestosave), unlist(rmse))
finalvips <- do.call(rbind, vips)

finalvips2 <- finalvips %>% group_by(Family) %>% dplyr::filter(Importance == max(Importance))
ggplot(finalvips2, mapping = aes(x = Variable)) + geom_bar() + my_theme() + labs(title = "xgboost Highest Variable Importance within each Family Model", x = "Variable", y = "Number of Families with Variable as Important")
ggsave(here("figures", "FamilyCountofVariableImportance.png"), width = 6, height = 4, unit = "in")
```

